Digitization of the Australian Parliamentary Debates (1901-2022)\footnote{Code is available at: https://github.com/lindsaykatz/hansard-proj}

### 2022-11-01 paper/paper.tex
paper files

---
title: "Digitization of the Australian Parliamentary Debates (1901-2022)^[Code is available at: https://github.com/lindsaykatz/hansard-proj]"
number-sections: true
format: 
  pdf:
    keep-tex: true
bibliography: references.bib
editor: visual
date: "now"
date-format: long
author:
  - Lindsay Katz^[University of Toronto, lindsay.katz@mail.utoronto.ca]
  - Rohan Alexander^[University of Toronto, rohan.alexander@utoronto.ca]
abstract: "Public knowledge of what is said in parliament is a key tenet of democracy. Parliamentary dialogue is also critical for political science research. In Australia, following the UK tradition, the written record of what is said in parliament is known as Hansard. While the Australian Hansard has always been publicly available, it has been difficult to use it for the purpose of large-scale macro and micro-level text analysis because it has not been available as a dataset of sufficient quality to be credibly analysed with statistical models. Following the lead of the LiPaD project which achieved this for Canada, our project aims to provide a new, comprehensive, high-quality database that captures all proceedings of the Australian parliamentary debates from 1901 to the present day using Hansard. Our dataset will be publicly available and linked to other datasets such as election results. The creation of this dataset will enable the exploration of questions that are not currently possible to explore, serving as a valuable resource for both researchers and policymakers. This work will provide a thorough description of the creation and computational underpinnings of this database, followed by a discussion of some applications. To this point, we have completed this for 2011 to 2022."
---

# Introduction {#sec-intro}

The official written record of parliamentary debates, formally known as Hansard, plays a fundamental role in capturing the history of political proceedings and facilitating the exploration of valuable research questions. Originating in the British parliament, the production of Hansard became tradition in a number of Commonwealth countries such as Canada, the United Kingdom, and Australia [@vice2017history]. Given the content and magnitude of these records, they have significance, particularly in the context of political science research. In the case of Canada, a team of researchers at the University of Toronto have digitized Hansard from 1901 to 2019, an endeavor called the Linked Parliamentary Data (LiPaD) project [@lipad]. Having a digitized version of Hansard enables researchers to perform advanced analyses on these records using text analysis tools and statistical modelling. Following the lead of the LiPaD project, in this paper we introduce a novel database for the Australian Hansard. This is composed of individual datasets for each sitting day in the House of Representatives, containing details on everything said in parliament in a form which can be readily used by researchers. With the development of tools for large-scale text analysis, this database will serve as a valuable resource for efficiently studying and exploring political behaviour in Australia over time.

The House of Representatives performs a number of crucial governmental functions, such as creating new laws and overseeing government expenditure [@australia2018house, ch. 1]. Sittings of the House follow a predefined order of business, regulated by procedural rules called standing orders [@australia2018house, ch. 8]. A typical sitting day in the Chamber has a number of scheduled proceedings including debates on government business, 90 second member statements, and Question Time [@australia2018house, ch. 8]. The Federation Chamber was created in 1994 as a subordinate debate venue of the Chamber. This allows for better time management of House business as its proceedings occur simultaneously with those of the Chamber [@australia2018house, ch. 21]. Sittings in the Federation Chamber are quite different to those of the Chamber in terms of their order of business and scope of discussion. Business matters discussed in the Federation Chamber are limited largely to intermediate stages of bill development, and the business of private Members [@australia2018house, ch. 21]. It is the recording and compilation of all of these proceedings on which Hansard is based, and it is essentially, but not entirely, verbatim.

A week or so after each sitting day, a transcript is available for download from the official Parliament of Australia website in both static PDF and extensible markup language (XML) form. The PDF is the official release, which is converted to typed text using Optical Character Recognition (OCR) technology [@sherratt]. This conversion allows for the production of the XML formatted transcript, though these are not always perfect conversions and are particularly flawed in the early decades of Hansard [@sherratt]. The PDF imposes formatting designed for humans to read with ease, whereas XML is designed for consistency and machine legibility. The nature of XML enables us to more easily use code to manipulate these records at scale, motivating our choice to develop our database solely using the XML formatted files. In cases where we were unsure on how to proceed with processing the XML, we defer first to the PDF, and then to the video recording of the proceeding if available.

The end goal for this work is to create a publicly available database of parsed Hansard transcripts from 1901 to present in a form which can be readily analyzed. More specifically, we aim to present each transcript in an ordered form with details on each speaker such as their political affiliation, the time stamp of their statement (where available), and whether they interjected (i.e. spoke out of turn). Thus far, this has been done for Hansard from mid-2011 to present.

We begin this paper with a description of what datasets for Australian Hansard already exist (@sec-existing). Next will be a detailed description of the contents of our database (@sec-description), followed by a thorough discussion of how this database was created (@sec-create).

# Existing Datasets {#sec-existing}

At present, the Hansard format that is available on the Parliament of Australia website is not accessible for large scale analysis. To this point, various researchers have had to create their own databases in an attempt to produce usable, complete Hansard data. For instance, @sherratt created an online, easy to read [database](http://historichansard.net) of Hansard from 1901 to 1980 using the XML files. These data can be navigated by year, parliament, people, and bills [@sherratt]. As well, @Leslie_2021_HoR_Aus has released data on all Members of the Australian House of Representatives from 1945 to 2019, and is developing an `R` package for these data. Further, @auspol created the `AustralianPoliticians` `R` package, which contains a number of datasets related to the political and biographical information of Australian federal politicians who were active between 1901 and 2021.

A number of papers exist which utilize components of Australian Hansard to explore various research topics. For example, @salisbury2011 used online Hansard records to explore the occurrences of unparliamentary comments by members of the House, where The Speaker tells that member to withdraw their remark. @rasiah2010framework worked with Question Time data from Hansard transcripts during February and March of 2003, to investigate resistance of politicians in answering questions about Iraq. In their work, @fraussen2018 use Hansard to quantify political prominence by investigating strategic mentions of interest groups by elected officials. Finally, the work of @alexander2021 offers a dataset of the Australian Hansard, along with an analysis of the effect of elections and changes in Prime Ministers upon topics mentioned in parliament. @alexander2021 created this database with the static PDF versions Hansard, using OCR to digitize these files into text which is suitable for analysis. This means there are considerable digitization errors especially in the first half of the dataset.

@@

# Description of the Database {#sec-description}

Our current database contains 670 comma-separated values (CSV) files, one for each sitting day of the House of Representatives from 10 May 2011 to 08 September 2022. We used two scripts to produce these files, which only have one subtle difference. Since the Federation Chamber was renamed in mid-2012 from the original title of Main Committee, the element under which all the Federation Chamber contents are nested in the XML was renamed as well [@australia2018house, ch. 21]. This difference in element name means that our original code for parsing Federation Chamber content would not work, as the specified node path would be incorrect. We accounted for this by creating a separate script with this name change. Our first script parses Hansard from 14 August 2012 to 08 September 2022, that being everything after the Federation Chamber name change, and the second script parses Hansard from 10 May 2011 to 28 June 2012. All output files in our database are formatted and processed in the exact same way. @fig-main15rows provides a snapshot of one of these files.

![First 15 rows of CSV file for Hansard on 25 February 2020](main_ex15.png){#fig-main15rows}

Each row of the CSV file contains an individual statement, with the unique name ID, electorate, and party information of the speaker. Of course, for general statements such as "Honourable members", these variables cannot be specified. Each proceeding in the Chamber begins with the business start, where the time that the Speaker took the chair is documented, as well as the acknowledgment of the country and reading of the prayers as governed by the standing orders of the House [@standingorders, ch. 6]. The Federation Chamber also begins with a business start, where it is stated at what time 5he Deputy Speaker takes the chair. For further clarification on each variable in the resulting CSV, @tbl-vars provides an overview of each variable in the database.

```{r}
#| warning: false
#| message: false
#| echo: false
#| label: tbl-vars
#| tbl-cap: "Summary and description of variables in our database"

# read in packages
library(readr)
library(tidyverse)
library(here)
library(XML)
library(lubridate)
library(kableExtra)

# read in example csv
main_ex <- read_csv("/Volumes/Verbatim/output/main/2020-02-25-main.csv", show_col_types = F)

main_ex <- main_ex %>% select(-c(gender, uniqueID))

# create tibble with var names and descriptions
main_ex_vars <- tibble("Variable" = names(main_ex),
       "Description" = c("Name of speaker",
                         "Row number",
                         "Speech number",
                         "Page number statement can be found on in official Hansard",
                         "Time of statement",
                         "Unique member identification code",
                         "Speaking member's electorate",
                         "Speaking member's party",
                         "Statement text",
                         "Flag for Federation Chamber (1 if Federation Chamber, 0 if Chamber)",
                         "Flag for sub-debate 1 contents (1 if sub-debate 1, 0 otherwise)",
                         "Flag for sub-debate 2 contents (1 if sub-debate 2, 0 otherwise)",
                         "Flag for question (1 if question, 0 otherwise)",
                         "Flag for answer (1 if answer, 0 otherwise)",
                         "Flag for question in writing (1 if question in writing, 0 otherwise)",
                         "Flag for interjection (1 if statement is an interjection, 0 otherwise)"))

# print tibble using kable
knitr::kable(main_ex_vars, booktabs=T) %>% 
  kable_styling(full_width = F, latex_options = "HOLD_position")
```

The `name`, `page.no`, `time.stamp`, `name.id`, `electorate`, `party`, and `body` variables all came directly from the XML contents. In addition to these variables, we added a number of flags to enable easy filtering of statements. For example, adding the `fedchamb_flag` provides a clear distinction between the proceedings of the Chamber with those of the Federation Chamber. As well, the `sub1_flag` and `sub2_flag` variables allow us to keep track of where various statements are being parsed from in the XML document. Further details on this XML structure will follow in @sec-create. The `question`, `answer`, and `q_in_writing` flags were added to identify statements belonging to Question Time, and the nature of these statements.[^1] We also added a flag for interjections, which are defined as statements made during one members speech by another member who is not the Speaker or Deputy Speaker. Further, the `speech_no` variable allows us to keep track of the speech number that each statement and interjection belongs to. For example, in @fig-main15rows, statements belonging to speech number 3 spans five rows, including two interjections. Having the speech number variable offers an easy way to group statements by speech or isolate specific speeches of interest. Finally, the `order` variable was added to keep track of the order in which everything was said, and was created after all statements and interjections were separated onto their own rows.

[^1]: More on this in @sec-qa

## Descriptive Statistics {#sec-stats}

A visual exploration of our database offers interesting insight into general characteristics of Hansard. First, we looked at the total number of sitting days in the House of Representatives for each year from 2000 to 2021. @fig-plot1 illustrates that the number of House sitting days tends to fluctuate over time, where within this time span the years 2013 and 2019 had the fewest sitting days (45), and 2014 had the most sitting days (76).

```{r}
#| warning: false
#| echo: false

# grab list of all file names
all_files <- list.files("/Volumes/Verbatim/input")

# extract year month and day
all_files <- all_files %>% 
  str_remove(".xml") %>% 
  as_tibble() %>% 
  mutate(year = as.numeric(str_extract(value, "^[[:digit:]]{4}")),
         month = as.numeric(str_extract(value, "(?<=\\-)[[:digit:]]{2}(?=\\-)")),
         day = as.numeric(str_extract(value, "(?<=\\-)[[:digit:]]{2}$")))

# plot of total number of sitting days per year
plot1 <- all_files %>% 
  filter(year!=2022) %>% 
  group_by(year) %>% 
  summarise(tot_days = n()) %>% 
  ggplot(aes(x=year, y=tot_days)) + 
  geom_point() +
  geom_line() + 
  theme_bw() +
  labs(x = "Year", 
       y = "Count")

# plot of number of sitting days per month from 2015-2020
# all_files %>% 
#   filter(year==2015|year==2016|year==2017|year==2018|year==2019|year==2020) %>%
#   mutate(month_abb = month.abb[month]) %>% 
#   group_by(year, month_abb) %>% 
#   summarise(avg_days = n()) %>% 
#   ggplot(aes(x=factor(month_abb, levels=month.abb), y=avg_days)) +
#   geom_col() + 
#   facet_wrap(~year) +
#   theme_bw() + 
#   labs(x = "Month", 
#        y="Count", 
#        title = "Number of sitting days per month (2015-2020)") +
#   theme(axis.text.x = element_text(angle=90, vjust=0.5, hjust=0.5))
```

```{r, out.width="80%"}
#| echo: false
#| message: false
#| fig-cap: "Total number of sitting days per year (2000-2021)"
#| label: fig-plot1

plot1
```

```{r}
#| echo: false
#| eval: false
# generating some descriptive statistics

# first grab list of all main CSVs
all_dfs <- list.files("/Volumes/Verbatim/output/main")

# define empty tibbles to store things in
interject_data <- tibble()
fedchamb_rows <- tibble()
unique_data <- tibble()

for(i in 1:length(all_dfs)){
  # define df
  this_df <- read_csv(paste0("/Volumes/Verbatim/output/main/", all_dfs[i]), show_col_types = F)

  # number of interjections
  this_interject_data <- this_df %>% group_by(interject) %>%
    summarise(n_interject=n()) %>%
    mutate(date = as.Date(str_extract(all_dfs[i], "\\d\\d\\d\\d\\-\\d\\d\\-\\d\\d"))) %>% 
    select(date, interject, n_interject)

  # number of rows in chamb and fedchamb
  this_fedchamb_rows <- this_df %>% group_by(fedchamb_flag) %>% 
    summarise(n_rows = n()) %>% 
    mutate(date = as.Date(str_extract(all_dfs[i], "\\d\\d\\d\\d\\-\\d\\d\\-\\d\\d"))) %>% 
    select(date, fedchamb_flag, n_rows)

  # number of unique things (NAs omitted) - need to fix a tiny bit more but better for now
  this_nameid <- this_df %>% select(name.id, fedchamb_flag) %>% 
    group_by(fedchamb_flag) %>% 
    unique() %>% 
    na.omit() %>% 
    summarise(n_name.id = n())

  this_electorate <- this_df %>% select(electorate, fedchamb_flag) %>% 
    group_by(fedchamb_flag) %>% 
    unique() %>% 
    na.omit() %>% 
    summarise(n_electorate = n())

  this_party <- this_df %>% select(party, fedchamb_flag) %>% 
    group_by(fedchamb_flag) %>% 
    unique() %>% 
    na.omit() %>% 
    summarise(n_party = n())

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  letterpaper,
  DIV=11,
  numbers=noendperiod]{scrartcl}

\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\KOMAoption{captions}{tableheading}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Digitization of the Australian Parliamentary Debates (1901-2022)},
  pdfauthor={Lindsay Katz; Rohan Alexander},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Digitization of the Australian Parliamentary Debates
(1901-2022)\footnote{Code is available at:
  https://github.com/lindsaykatz/hansard-proj}}
\author{Lindsay Katz\footnote{University of Toronto,
  lindsay.katz@mail.utoronto.ca} \and Rohan Alexander\footnote{University
  of Toronto, rohan.alexander@utoronto.ca}}
\date{October 27, 2022}

\begin{document}
\maketitle
\begin{abstract}
Public knowledge of what is said in parliament is a key tenet of
democracy. Parliamentary dialogue is also critical for political science
research. In Australia, following the UK tradition, the written record
of what is said in parliament is known as Hansard. While the Australian
Hansard has always been publicly available, it has been difficult to use
it for the purpose of large-scale macro and micro-level text analysis
because it has not been available as a dataset of sufficient quality to
be credibly analysed with statistical models. Following the lead of the
LiPaD project which achieved this for Canada, our project aims to
provide a new, comprehensive, high-quality database that captures all
proceedings of the Australian parliamentary debates from 1901 to the
present day using Hansard. Our dataset will be publicly available and
linked to other datasets such as election results. The creation of this
dataset will enable the exploration of questions that are not currently
possible to explore, serving as a valuable resource for both researchers
and policymakers. This work will provide a thorough description of the
creation and computational underpinnings of this database, followed by a
discussion of some applications. To this point, we have completed this
for 2011 to 2022.
\end{abstract}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, interior hidden, sharp corners, breakable, frame hidden, enhanced, boxrule=0pt]}{\end{tcolorbox}}\fi

\hypertarget{sec-intro}{%
\section{Introduction}\label{sec-intro}}

The official written record of parliamentary debates, formally known as
Hansard, plays a fundamental role in capturing the history of political
proceedings and facilitating the exploration of valuable research
questions. Originating in the British parliament, the production of
Hansard became tradition in a number of Commonwealth countries such as
Canada, the United Kingdom, and Australia (Vice and Farrell 2017). Given
the content and magnitude of these records, they have significance,
particularly in the context of political science research. In the case
of Canada, a team of researchers at the University of Toronto have
digitized Hansard from 1901 to 2019, an endeavor called the Linked
Parliamentary Data (LiPaD) project (Beelen et al. 2017). Having a
digitized version of Hansard enables researchers to perform advanced
analyses on these records using text analysis tools and statistical
modelling. Following the lead of the LiPaD project, in this paper we
introduce a novel database for the Australian Hansard. This is composed
of individual datasets for each sitting day in the House of
Representatives, containing details on everything said in parliament in
a form which can be readily used by researchers. With the development of
tools for large-scale text analysis, this database will serve as a
valuable resource for efficiently studying and exploring political
behaviour in Australia over time.

The House of Representatives performs a number of crucial governmental
functions, such as creating new laws and overseeing government
expenditure (House of Representatives 2018, ch.~1). Sittings of the
House follow a predefined order of business, regulated by procedural
rules called standing orders (House of Representatives 2018, ch.~8). A
typical sitting day in the Chamber has a number of scheduled proceedings
including debates on government business, 90 second member statements,
and Question Time (House of Representatives 2018, ch.~8). The Federation
Chamber was created in 1994 as a subordinate debate venue of the
Chamber. This allows for better time management of House business as its
proceedings occur simultaneously with those of the Chamber (House of
Representatives 2018, ch.~21). Sittings in the Federation Chamber are
quite different to those of the Chamber in terms of their order of
business and scope of discussion. Business matters discussed in the
Federation Chamber are limited largely to intermediate stages of bill
development, and the business of private Members (House of
Representatives 2018, ch.~21). It is the recording and compilation of
all of these proceedings on which Hansard is based, and it is
essentially, but not entirely, verbatim.

A week or so after each sitting day, a transcript is available for
download from the official Parliament of Australia website in both
static PDF and extensible markup language (XML) form. The PDF is the
official release, which is converted to typed text using Optical
Character Recognition (OCR) technology (Sherratt 2016). This conversion
allows for the production of the XML formatted transcript, though these
are not always perfect conversions and are particularly flawed in the
early decades of Hansard (Sherratt 2016). The PDF imposes formatting
designed for humans to read with ease, whereas XML is designed for
consistency and machine legibility. The nature of XML enables us to more
easily use code to manipulate these records at scale, motivating our
choice to develop our database solely using the XML formatted files. In
cases where we were unsure on how to proceed with processing the XML, we
defer first to the PDF, and then to the video recording of the
proceeding if available.

The end goal for this work is to create a publicly available database of
parsed Hansard transcripts from 1901 to present in a form which can be
readily analyzed. More specifically, we aim to present each transcript
in an ordered form with details on each speaker such as their political
affiliation, the time stamp of their statement (where available), and
whether they interjected (i.e.~spoke out of turn). Thus far, this has
been done for Hansard from mid-2011 to present.

We begin this paper with a description of what datasets for Australian
Hansard already exist (Section~\ref{sec-existing}). Next will be a
detailed description of the contents of our database
(Section~\ref{sec-description}), followed by a thorough discussion of
how this database was created (Section~\ref{sec-create}).

\hypertarget{sec-existing}{%
\section{Existing Datasets}\label{sec-existing}}

At present, the Hansard format that is available on the Parliament of
Australia website is not accessible for large scale analysis. To this
point, various researchers have had to create their own databases in an
attempt to produce usable, complete Hansard data. For instance, Sherratt
(2016) created an online, easy to read
\href{http://historichansard.net}{database} of Hansard from 1901 to 1980
using the XML files. These data can be navigated by year, parliament,
people, and bills (Sherratt 2016). As well, Leslie (2021) has released
data on all Members of the Australian House of Representatives from 1945
to 2019, and is developing an \texttt{R} package for these data.
Further, Alexander and Hodgetts (2021) created the
\texttt{AustralianPoliticians} \texttt{R} package, which contains a
number of datasets related to the political and biographical information
of Australian federal politicians who were active between 1901 and 2021.

A number of papers exist which utilize components of Australian Hansard
to explore various research topics. For example, Salisbury (2011) used
online Hansard records to explore the occurrences of unparliamentary
comments by members of the House, where The Speaker tells that member to
withdraw their remark. Rasiah (2010) worked with Question Time data from
Hansard transcripts during February and March of 2003, to investigate
resistance of politicians in answering questions about Iraq. In their
work, Fraussen, Graham, and Halpin (2018) use Hansard to quantify
political prominence by investigating strategic mentions of interest
groups by elected officials. Finally, the work of Alexander and
Alexander (2021) offers a dataset of the Australian Hansard, along with
an analysis of the effect of elections and changes in Prime Ministers
upon topics mentioned in parliament. Alexander and Alexander (2021)
created this database with the static PDF versions Hansard, using OCR to
digitize these files into text which is suitable for analysis. This
means there are considerable digitization errors especially in the first
half of the dataset.

@@

\hypertarget{sec-description}{%
\section{Description of the Database}\label{sec-description}}

Our current database contains 670 comma-separated values (CSV) files,
one for each sitting day of the House of Representatives from 10 May
2011 to 08 September 2022. We used two scripts to produce these files,
which only have one subtle difference. Since the Federation Chamber was
renamed in mid-2012 from the original title of Main Committee, the
element under which all the Federation Chamber contents are nested in
the XML was renamed as well (House of Representatives 2018, ch.~21).
This difference in element name means that our original code for parsing
Federation Chamber content would not work, as the specified node path
would be incorrect. We accounted for this by creating a separate script
with this name change. Our first script parses Hansard from 14 August
2012 to 08 September 2022, that being everything after the Federation
Chamber name change, and the second script parses Hansard from 10 May
2011 to 28 June 2012. All output files in our database are formatted and
processed in the exact same way. Figure~\ref{fig-main15rows} provides a
snapshot of one of these files.

\begin{figure}

{\centering \includegraphics{main_ex15.png}

}

\caption{\label{fig-main15rows}First 15 rows of CSV file for Hansard on
25 February 2020}

\end{figure}

Each row of the CSV file contains an individual statement, with the
unique name ID, electorate, and party information of the speaker. Of
course, for general statements such as ``Honourable members'', these
variables cannot be specified. Each proceeding in the Chamber begins
with the business start, where the time that the Speaker took the chair
is documented, as well as the acknowledgment of the country and reading
of the prayers as governed by the standing orders of the House
(Department of the House of Representatives 2019, ch.~6). The Federation
Chamber also begins with a business start, where it is stated at what
time 5he Deputy Speaker takes the chair. For further clarification on
each variable in the resulting CSV, Table~\ref{tbl-vars} provides an
overview of each variable in the database.

\hypertarget{tbl-vars}{}
\begin{table}[H]
\caption{\label{tbl-vars}Summary and description of variables in our database }\tabularnewline

\centering
\begin{tabular}{ll}
\toprule
Variable & Description\\
\midrule
name & Name of speaker\\
order & Row number\\
speech\_no & Speech number\\
page.no & Page number statement can be found on in official Hansard\\
time.stamp & Time of statement\\
\addlinespace
name.id & Unique member identification code\\
electorate & Speaking member's electorate\\
party & Speaking member's party\\
body & Statement text\\
fedchamb\_flag & Flag for Federation Chamber (1 if Federation Chamber, 0 if Chamber)\\
\addlinespace
sub1\_flag & Flag for sub-debate 1 contents (1 if sub-debate 1, 0 otherwise)\\
sub2\_flag & Flag for sub-debate 2 contents (1 if sub-debate 2, 0 otherwise)\\
question & Flag for question (1 if question, 0 otherwise)\\
answer & Flag for answer (1 if answer, 0 otherwise)\\
q\_in\_writing & Flag for question in writing (1 if question in writing, 0 otherwise)\\
\addlinespace
interject & Flag for interjection (1 if statement is an interjection, 0 otherwise)\\
\bottomrule
\end{tabular}
\end{table}

The \texttt{name}, \texttt{page.no}, \texttt{time.stamp},
\texttt{name.id}, \texttt{electorate}, \texttt{party}, and \texttt{body}
variables all came directly from the XML contents. In addition to these
variables, we added a number of flags to enable easy filtering of
statements. For example, adding the \texttt{fedchamb\_flag} provides a
clear distinction between the proceedings of the Chamber with those of
the Federation Chamber. As well, the \texttt{sub1\_flag} and
\texttt{sub2\_flag} variables allow us to keep track of where various
statements are being parsed from in the XML document. Further details on
this XML structure will follow in Section~\ref{sec-create}. The
\texttt{question}, \texttt{answer}, and \texttt{q\_in\_writing} flags
were added to identify statements belonging to Question Time, and the
nature of these statements.\footnote{More on this in
  Section~\ref{sec-qa}} We also added a flag for interjections, which
are defined as statements made during one members speech by another
member who is not the Speaker or Deputy Speaker. Further, the
\texttt{speech\_no} variable allows us to keep track of the speech
number that each statement and interjection belongs to. For example, in
Figure~\ref{fig-main15rows}, statements belonging to speech number 3
spans five rows, including two interjections. Having the speech number
variable offers an easy way to group statements by speech or isolate
specific speeches of interest. Finally, the \texttt{order} variable was
added to keep track of the order in which everything was said, and was
created after all statements and interjections were separated onto their
own rows.

\hypertarget{sec-stats}{%
\subsection{Descriptive Statistics}\label{sec-stats}}

A visual exploration of our database offers interesting insight into
general characteristics of Hansard. First, we looked at the total number
of sitting days in the House of Representatives for each year from 2000
to 2021. Figure~\ref{fig-plot1} illustrates that the number of House
sitting days tends to fluctuate over time, where within this time span
the years 2013 and 2019 had the fewest sitting days (45), and 2014 had
the most sitting days (76).

\begin{figure}

{\centering \includegraphics[width=0.8\textwidth,height=\textheight]{paper_files/figure-pdf/fig-plot1-1.pdf}

}

\caption{\label{fig-plot1}Total number of sitting days per year
(2000-2021)}

\end{figure}

In Figure~\ref{fig-plot2}, we consider the average number of rows per
sitting day per year, disaggregated by Chamber and Federation Chamber
proceedings. We can see that 2014 has the highest average number of rows
per day, and 2021 has the lowest average number of rows per day. There
appears to be less fluctuation in the average number of rows per day in
the Federation Chamber than there is in that of the Chamber.

\begin{figure}

{\centering \includegraphics[width=0.8\textwidth,height=\textheight]{paper_files/figure-pdf/fig-plot2-1.pdf}

}

\caption{\label{fig-plot2}Average number of rows in Hansard per year, by
Chamber and Federation Chamber}

\end{figure}

Next we were interested in the percentage of rows in each Hansard which
are flagged as interjections, using our definition of interjections
mentioned above. We found that on average, approximately 18\% of rows
are characterized as interjections. Further, of all sitting days in our
database, 25 June 2014 had the greatest percentage of rows being
interjections, which was approximately 33\%. The day with the smallest
percentage of interjections was 23 March 2015, at approximately 3\%.
Based on Figure~\ref{fig-plot3}, there do not appear to be any
significant fluctuations or patterns in the percentage of rows that are
interjections across our time span.

\begin{figure}

{\centering \includegraphics[width=0.8\textwidth,height=\textheight]{paper_files/figure-pdf/fig-plot3-1.pdf}

}

\caption{\label{fig-plot3}Percentage of rows in Hansard that are
interjections, 10 May 2011 - 08 September 2022}

\end{figure}

Finally, Figure~\ref{fig-unique_plots} summarizes the number of unique
speakers, electorates, and parties found in Hansard over time. In all
three panels, the data exhibit the same general pattern over time for
both the House of Representatives and the Federation Chamber, where the
magnitude for the House data is consistently greater than that of the
Federation Chamber. Panels A and B in Figure~\ref{fig-unique_plots}
illustrate that the number of unique names and electorates follow a very
consistent trend across the entire time frame. Panel C shows that the
number of unique parties has increased slightly since 2018.

\begin{figure}

{\centering \includegraphics{unique_plots.png}

}

\caption{\label{fig-unique_plots}Number of unique names (A), electorates
(B), and parties (C) found on each sitting day.}

\end{figure}

\begin{table}

\end{table}

\hypertarget{sec-create}{%
\section{Creation of the Database}\label{sec-create}}
